# Wind use in foraging trips

Demonstrating the use of winds during a foraging trip by streaked shearwaters nesting in Funakoshi-Ohshima island, Japan. To do this, we use a combination foraging estimates (generated by the methods described in this [2018 study](https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0254454)) and wind vector estimates (generated by the methods described in this [2017 study](https://www.science.org/doi/10.1126/sciadv.1700097)).

Once we read in the data, we can plot one example of the bird's tracks, foraging point (FP), and estimated wind vectors:

```{r, echo=FALSE}
library(dplyr)
library(plyr)
library(ggplot2)
library(sf)
library(rnaturalearth)
library(sp)
library(RColorBrewer)
library(reshape)
library(circular)
library(CircStats)
library(CircMLE)
library(extrafont)
# LOAD FORAGING DATA
if(Sys.info()['sysname'] == "Darwin"){
    # load("/Volumes/GoogleDrive-112399531131798335686/My Drive/PhD/Data/2019Shearwater/2019Dat.RData")
    load("/Volumes/GoogleDrive-112399531131798335686/My Drive/PhD/Data/DatEth2018.RData")
    load("/Volumes/GoogleDrive-112399531131798335686/My Drive/PhD/Data/DatEth2019.RData")
    outloc <- "/Volumes/GoogleDrive-112399531131798335686/My Drive/PhD/Manuscripts/BehaviourIdentification/Figures/"
} else {
    # load("E:/My Drive/PhD/Data/2019Shearwater/2019Dat.RData")
    load("E:/My Drive/PhD/Data/DatEth2018.RData")
    load("E:/My Drive/PhD/Data/DatEth2019.RData")
    outloc <- "E:/My Drive/PhD/Manuscripts/BehaviourIdentification/Figures/"
}
D18 <- bind_rows(Dat)
D19 <- bind_rows(Dat19)
allD <- data.frame(DT=c(D18$DT, D19$DT),
    lat = c(D18$Lat, D19$Lat),
    lon = c(D18$Lon, D19$Lon),
    Forage = c(D18$Forage, D19$Forage),
    tagID = c(D18$tagID, D19$tagID),
    Day = c(D18$Day, D19$Day),
    Sex = c(D18$Sex, D19$Sex),
    distTrav = c(D18$recalDist, D19$recalDist),
    spTrav = c(D18$spTrav, D19$spTrav),
    recalSp = c(D18$recalSp, D19$recalSp),
    distFk = c(D18$distFromFk, D19$distFromFk),
    tripN = c(D18$tripN, D19$tripN),
    tripL = c(D18$tripL, D19$tripL),
    tkb = c(D18$tkb, D19$tkb),
    dv = c(D18$dv, D19$dv),
    UTME = c(D18$UTME, D19$UTME),
    UTMN = c(D18$UTMN, D19$UTMN))
allD$Year <- format(allD$DT, format = "%Y")
allD$forage <- allD$dv == 1 | allD$tkb == 1
allD$yrID <- paste(format(allD$DT,'%Y'),"_",sub("\\_S.*","",allD$tagID),sep="")
# LOAD WIND DATA
if(Sys.info()['sysname'] == "Darwin"){
  load("/Volumes/GoogleDrive-112399531131798335686/My Drive/PhD/Data/WindCalculations1819.RData")
} else {
  load("E:/My Drive/PhD/Data/WindCalculations1819.RData")
}
```

```{r, echo=FALSE}
japan <- ne_countries(scale = "medium", country = "Japan", returnclass = "sf")
WindDat$WSpd <- sqrt(WindDat$X^2 + WindDat$Y^2)
yrids <- unique(WindDat$yrID)
ggplot() + geom_sf(data = japan, fill = '#969696', colour = '#969696') +
    geom_path(data = allD[allD$yrID == yrids[5],], aes(x = lon, y = lat)) +
    coord_sf(xlim = c(140, 146), ylim = c(39, 44)) +
    geom_point(data = allD[allD$yrID == yrids[5] & allD$forage == 1,], aes(x = lon, y = lat), pch = 21, fill = "deepskyblue") + # nolint
    geom_spoke(data = WindDat[WindDat$yrID == yrids[5],], aes(x = lon, y = lat, colour = WSpd, angle = WHead), arrow = arrow(length = unit(0.05,"inches")),
    radius = .5*(WindDat$WSpd[WindDat$yrID == yrids[5]]/max(WindDat$WSpd[WindDat$yrID == yrids[5]]))) +
    scale_colour_distiller(name="Wind Speed (m/s)", direction = 1, palette = "YlOrRd") +
    scale_y_continuous(breaks = c(39,40,41,42,43,44), labels = c("39","40","41","42","43","44"), name = paste("Latitude (","\u00b0N",")", sep = "")) +
    scale_x_continuous(breaks = c(140,142,144,146),labels = c("140", "142", "144", "146"), name = paste("Longitude (","\u00b0E",")", sep = "")) +
    theme_bw() + theme(panel.grid = element_blank()) +
    theme(panel.border = element_rect(colour = 'black', fill = NA), text = element_text(face="plain",size = 12), axis.text = element_text(size = 12))
```

### The next foraging point

The aim of this analysis is to gather and examine the wind conditions in which shearwaters flew in as they approached FPs. Procellariiforms, including shearwaters, are believed to typically use olfaction during foraging trips to identify potential FPs and direct their headings. Therefore, for each wind estimate, we must find the closest FP after each wind estimate for the same tag. We also add the speed travelled at and the trip length.

### Binned data

To examine the data, we bin each wind value by it's distance from the next FP. In the following example, we bin for every 10 km, then plot the density (kernel density estimate) of relative wind headings (the wind heading relative to the bird's movement) for each bin (less than 90km to the next FP)

```{r, echo=FALSE}
breaks<-seq(from=0,to=round_any(max(WindDat$distTo,na.rm=T),10,f=ceiling),by=10)
WindDat$bin10 <- cut(WindDat$distTo, breaks = breaks, include.lowest=T,right=F)
mnW <- ddply(WindDat, "bin10", summarise, grp.mean=mean(aligned))
# Cairo(width=15, height = 15, file = paste(figLoc,"DistRelDensity.svg",sep=""),type="svg", bg = "transparent", dpi = 300, units="in")
bin10ns <- WindDat[WindDat$distTo < 90,] %>% group_by(bin10) %>% dplyr::summarise(length(unique(yrID)))
ggplot(WindDat[WindDat$distTo > 0 &WindDat$distTo < 90,], aes(x = aligned, colour = bin10)) +#max(WindDat$distTo),], aes(x = aligned, colour = bin10)) +
  # geom_histogram(alpha=.2,fill=NA,position="dodge")
  geom_density(alpha=.5,show.legend=FALSE)+stat_density(aes(x=aligned, colour=bin10), size=1.1, geom="line",position="identity") +
  scale_x_continuous(name = "Relative wind heading", breaks=c(-pi, -pi/2, 0, pi/2, pi), labels=c("Tail","Side","Head","Side","Tail")) + ylab("") + theme_bw() + theme(panel.grid = element_blank()) +
  scale_colour_manual(name="Distance to next \nforaging spot (km)", values = rev(brewer.pal(9,"Blues")),
    labels=paste(gsub(",",":",gsub('[[)]',"",sort(unique(WindDat$bin10[WindDat$distTo < 90])))),", (", as.character(unlist(bin10ns[,2])),")",sep="")) +
  theme(panel.border = element_rect(colour = 'black', fill = NA), text = element_text(size = 12), axis.text = element_text(size = 12)) + 
  scale_y_continuous(name="Proportion across all birds (%)", breaks=seq(0,0.6,.1),labels=seq(0,60,10))
```

We can see there appears to be a strong bimodal distribution from 90 to 40 km to the next FP, changing to birds flying in more headwind conditions once within 20 km to the next FP. This distance is in good agreement with other studies examining olfactory and visual reliance on foraging behaviour in seabirds.